---
title: "Untitled"
author: "HAZIM FITRI"
date: "2025-06-22"
output: html_document
---

# 1. Image Data Analysis

Pixel (aka picture element in scientific term) has intensity.

plane origin (0,0) start at top left instead of bottom left

-   Binary images: black (0) and white (1) only
-   Gray scale intensity (0, 255) lower darker (0), higher lighter/brighter(255).
-   Color image: combination of each **Red**, **Green**, and **Blue** and their intensity (0, 255)

## 1.2 Morphological Processing

-   Dilated (+) pixel boundary

-   Eroded (-) pixel boundary

```{r}
library(imager)

file = system.file('extdata/parrots.png',package='imager')
parrots = load.image(file)
plot(parrots)
```

```{r}
par(mfrow=c(2,2))
parrots.blurry = isoblur(parrots,10) #Blurry parrots
parrots.xedges = deriche(parrots,2,order=2,axis="x") #Edge detector along x-axis
parrots.yedges = deriche(parrots,2,order=2,axis="y") #Edge detector along y-axis
gray.parrots = grayscale(parrots) #Make grayscale
plot(parrots.blurry)
plot(parrots.xedges)
plot(parrots.yedges)
plot(gray.parrots)
```

```{r}
rose <- load.image("Rose.png") #Load image
save.image(rose,"Rose.jpeg") #save image
plot(rose) #Rose
```

```{r denoising}
# Denoising
# create a noise image
par(mfrow=c(1,2))
birds = load.image(system.file('extdata/Leonardo_Birds.jpg',package='imager'))
birds.noisy = (birds + .5 * rnorm(prod(dim(birds))))
plot(birds, main = 'Original')
plot(birds.noisy, main = 'Noised Image')
```

```{r}
layout(t(1:2))
plot(birds.noisy,main="Original")
isoblur(birds.noisy,5) %>% 
  plot(main="Blurred")
```

```{r}
layout(t(1:2))
plot(birds.noisy,main="Original")
blur_anisotropic(birds.noisy, ampl=1e3, sharp=.3) %>% 
plot(main="Blurred (anisotropic)")
```

```{r}
 #Resize, rotate, ext.
thmb = resize(parrots,round(width(parrots)/10),round(height(parrots)/10))
plot(thmb,main="Thumbnail") # Pixellated parrots
```

## 1.3 Image Transformation

```{r}
par(mfrow = c(2, 2))
resize(parrots,-10,-10) %>% plot(main = 'Resize (smaller)')
imrotate(parrots,30) %>% plot(main="Rotating") #Rotate
imshift(parrots,40,20) %>% plot(main="Shifting") #Shift
imshift(parrots,100,100,boundary=2) %>% plot(main="Shifting (circular)") #Shift circular
```

## 1.4 Histogram Equalizations

```{r}
#Histogram equalisations
par(mfrow = c(2,2))
plot(boats)
grayscale(boats) %>% hist(main="Luminance values in boats picture")
R(boats) %>% hist(main="Red channel values in boats picture")

library(ggplot2)
library(dplyr)
bdf <- as.data.frame(boats)
bdf <- mutate(bdf, channel = factor(cc, labels = c('R','G','B')))
ggplot(bdf,aes(value,col=channel))+geom_histogram(bins=30)+facet_wrap(~ channel)
```

```{r}
#Normalizing
x = rnorm(100)
layout(t(1:2))
hist(x, main = "Histogram of x")
f = ecdf(x)
hist(f(x), main = "Histogram of ecdf(x)")

boats.g = grayscale(boats)
f = ecdf(boats.g)
plot(f, main = "Empirical CDF of luminance values")
f(boats.g) %>% hist(main = "Transformed luminance values")
f(boats.g) %>% str
f(boats.g) %>% 
  as.cimg(dim = dim(boats.g)) %>% 
  plot(main = "With histogram equalisation")
```

### 🔢 **Step-by-step Manual Histogram Equalization**

Let's say the grayscale image has intensity values ranging from 0 to 7 (i.e., 3-bit image, L = 8). These are the steps described in the file:

#### **Step 1: Compute the Frequency Table (Histogram)**

Count how many pixels have each intensity value:

| Gray Level (i) | Frequency (f(i)) |
|----------------|------------------|
| 0              | 4                |
| 1              | 8                |
| 2              | 10               |
| ...            | ...              |

#### **Step 2: Compute PDF (Probability Distribution Function)**

PDF is calculated by dividing the frequency by the total number of pixels (N):

PDF(i)=f(i)N\\text{PDF}(i) = \\frac{f(i)}{N}

#### **Step 3: Compute CDF (Cumulative Distribution Function)**

The CDF is computed as:

CDF(i)=∑j=0iPDF(j)\\text{CDF}(i) = \\sum\_{j=0}\^{i} \\text{PDF}(j)

#### **Step 4: Multiply CDF by (L-1) and Round Off**

L is the number of gray levels (usually 256 for 8-bit, so L - 1 = 255).

New Intensity(i)=round[(L−1)×CDF(i)]\\text{New Intensity}(i) = \\text{round}[(L - 1) \\times \\text{CDF}(i)]

#### **Step 5: Replace Old Pixel Values**

For each pixel in the image, replace the original intensity with the newly computed one from Step 4.

### 📌 **Example**

Assume L = 8 and total pixels N = 25. Gray levels range from 0 to 7.

| Gray Level | Frequency (f) | PDF = f/N | CDF  | New Value = round(CDF × 7) |
|------------|---------------|-----------|------|----------------------------|
| 0          | 2             | 0.08      | 0.08 | 1                          |
| 1          | 3             | 0.12      | 0.20 | 1                          |
| 2          | 4             | 0.16      | 0.36 | 3                          |
| 3          | 3             | 0.12      | 0.48 | 3                          |
| 4          | 5             | 0.20      | 0.68 | 5                          |
| 5          | 2             | 0.08      | 0.76 | 5                          |
| 6          | 3             | 0.12      | 0.88 | 6                          |
| 7          | 3             | 0.12      | 1.00 | 7                          |

### 📈 Plotting

-    **Original Histogram**: bar plot of original intensity values

-    **Equalized Histogram**: bar plot of new intensity values

-   Would you like an R script to apply this manually to an actual image?

## 1.5 Morphological Operations

Used to analyze and manipulate the structure or form of objects in an image.

-   Erosion : Shrink image by removing pixels on boundaries

-   Dilation : Expands objects by adding pixels to boundaries

Used for :

1.  Removing noise
2.  Filling gaps
3.  Finding boundaries
4.  Preparing images for further analysis like object detection or classification.

```{r}
#Morphological operations
im <- load.example("birds")
plot(im)

im.g <- grayscale(im)
plot(im.g)

layout(t(1:3))
threshold(im.g,"20%") %>% plot
threshold(im.g,"15%") %>% plot
threshold(im.g,"10%") %>% plot

```

## 1.X Segmentation

```{r}
plot(parrots, main = 'Original Image')

# convert to grayscale
im.g = grayscale(parrots)
plot(im.g, main = 'Grayscale Image')

# thresholding (top 30% intensity = white and vice versa)
im.b = threshold(im.g, "30%")
plot(im.b, main = "Binary Image")

# remove small object
im.c <- erode_square(im.b, 3)
im.c <- dilate_square(im.c, 3)
plot(im.c, main = "Cleaned Binary Image")

im.labeled <- label(im.c)
plot(im.labeled, main = "Segmented Regions (Labeled)")

num_segments <- max(im.labeled)
cat("Total segments found:", num_segments, "\n")
seg1 <- im.labeled == 1
plot(seg1, main = "Segment 1")

# Generate a color for each segment
num_segments <- max(im.labeled)
colors <- rainbow(num_segments)  # You can use other palettes too

# Convert label to numeric
px <- as.cimg(im.labeled)

# Apply color
colored <- colorise(im = im.labeled, px = px, col = colors)
plot(colored, main = "Colored Segmented Regions")
```

## 1.6 Splitting and concatenating images

```{r}
imsplit(parrots, "x", 4) %>% plot

```

### 1.7 Morphological Operations

#### 1.7.1 Thresholding, and correcting for varying luminance

```{r morphology}
im = load.image(system.file('extdata/Leonardo_Birds.jpg',package='imager'))
im.g = grayscale(im)
df = as.data.frame(im.g)
m = lm(value ~ x + y, data = df)
im.f = im.g - fitted(m)
plot(im.f, main = 'removed trend')
```

```{r}
im = load.image(system.file('extdata/Leonardo_Birds.jpg',package='imager'))
im.g = grayscale(im)
df = as.data.frame(im.g)
m = lm(value ~ x + y, data = df)
im.f = im.g - fitted(m)
im.t = threshold(im.f, "16%")
px = as.pixset(1 - im.t)
shrink(px, 3) %>% grow(3) %>% plot(main = 'shrink, then grow')
```

# 2. Audio Data Analysis

## 2.1 Fourier Transform and Spectogram

-   Pitching = frequency (kenyaringan suara)

-   volume = amplitude (kekuatan suara)

```{r}
library(tuneR)
library(seewave)

sr = 8000  # sampling rate
t = seq(0, 2, 1/sr)  # time vector for 2 seconds
y = (2^15 - 1) * sin(2 * pi * 440 * t)  # 440 Hz sine wave

y = as.integer(y)  # convert to 16-bit integers

plot(y[1:500], type = "l")  # plot the wave

w = Wave(left = y, samp.rate = sr, bit = 16)  # create Wave object
# play(w)  # play the sound
```

```{r}
library(tuneR)
library(seewave)

# Let's make a sine wave and play it
help(Wave)
sr = 8000  # the sampling rate
t = seq(0, 2, 1/sr)  # times in secs if sample for 2 seconds at 8KHz
y = (2 ^ 15 - 1) * sin(2 * pi * 440 * t)  # sine wave at 440 Hz scaled to fill out 16 bit range
plot(y[1:500], type = "l")
w = Wave(y, samp.rate = sr, bit = 16)  # make the Wave representation
#play(w)

plot(y[1:500], type="l")
y1 = (2 ^ 15 - 1) * sin(2 * pi * 220 * t)
w1 = Wave(y1, samp.rate = sr, bit = 16)
#play(w1)

# Compare w and w1
par(mfrow = c(2, 1))  # layout parallel
plot(y[1:500], type = "l")
plot(y1[1:500], type = "l")

# Compare in 1 layout (make redundant graph for easier comparison)
par(mfrow = c(1, 1))
plot(y[1:500], type = "l")  # plot y above
lines(y1[1:500], lty = 3, col = "red")  # plot y1 in plot y

# Try create another
wsum = normalize(w + w1, unit = '16')  # must normalize to make it also 16hz
play(wsum)
plot((y + y1)[1:500], type = "l")  # plot from 1 to 500

w3 = normalize(bind(w, w1, wsum), unit = '16')
play(w3)

y3 <- c(y, y1, y+y1)
plot(y3, type="l")

# The above to create manually. Next we will use the function in tuneR
# Other types of waveforms
Au1 <- sine(500, duration=100000)  # 100=freq, duration 100000 = ~2.2s
play(Au1)
writeWave(Au1, 'Audio1.wav')  # save data
Au11 <- readWave('Audio1.wav')  # read the audio

Au2 <- noise(duration = 100000)  # look noise function from ?sine
play(Au2)  # seems sound broken like TV/radio
Au2 <- noise(kind='pink', duration = 100000)  # change to pink noise (correlated noise)
play(Au2)

Au3 <- pulse(220, duration = 100000)
play(Au3)

# Plot 3 signals above
par(mfrow=c(3,1))
plot(Au1[1:1000])  # The first 1000 signal for Au1
plot(Au2[1:1000])
plot(Au3[1:1000])

# Other signals in ?sine
Au4 <- sawtooth(100, duration = 100000)
play(Au4)
plot(Au4[1:1000])
plot(Au4[1:2000])

Au5 <- square(200, duration = 100000)
play(Au5)
plot(Au5[1:2000])

# Combine all Au1 to Au5
Asum = normalize(bind(Au1, Au2, Au3, Au4, Au5), unit='32')
play(Asum)

Au6 = normalize(bind(Au1 + Au2), unit='32')
play(Au6)

```

```{r}
sr = 8000
t = seq(0, 2, 1/sr)
y = (2^15-1)*sin(2*pi*440*t)
plot(y[1:500], type="l")
```

```{r}
spectrum(y) # spectrum(y, span=20, log=c("no"))
```

## 2.2 Plot Frequency Spectrum

```{r plot_frequency_spectrum}
plot.frequency.spectrum <- function(X.k, xlimits=c(0,length(X.k))) {
  plot.data <- cbind(0:(length(X.k)-1), Mod(X.k))

  # TODO: why this scaling is necessary?
  plot.data[2:length(X.k),2] <- 2*plot.data[2:length(X.k),2]

  plot(plot.data, t="h", lwd=2, main="",
       xlab="Frequency (Hz)", ylab="Strength",
       xlim=xlimits, ylim=c(0,max(Mod(plot.data[,2]))))
}
```

## 2.3 Fast Fourier Transform

```{r}
par(mfrow=c(1,3))
Yk <- fft(y)
plot.frequency.spectrum(Yk)
plot.frequency.spectrum(Yk[1:5000])  # look roughly where the signal is
plot.frequency.spectrum(Yk[1:1000])  # around 800+
```

## 2.4 Timer

```{r}
# data tico from notes seewave
data(tico)
# play(tico)
par(mfrow=c(1,2))
timer(tico, f=22050, threshold=5, msmooth=c(50,0))  # f is freq, msmooth to smooth the line
timer(tico, f=22050, threshold=5, msmooth=c(100,0))
```

```{r}
plot(tico)
spectrum(tico)
```

```{r}
# audio file have left channel and right channel, so we want only left channel
Ticok = fft(tico@left)  
plot.frequency.spectrum(Ticok[1:(length(Ticok)/2)])  # mirror, focus only one side
```

```{r}
dynspec(tico, wl=1024, osc=T)  # function seewave, ocs = T (time)
```

```{r}
spectro(tico)  # give spectrogram
```

```{r}
meanspec(tico)  # obtain average for the whole time length
```

### Baby cry

```{r}
z = readWave("babycry.wav")
# play(z)
timer(z, f=22050, threshold=30, msmooth=c(300,0))
```

```{r}
plot(z)
```

```{r}
par(mfrow=c(1,2))
Zk <- fft(z@left)
plot.frequency.spectrum(Zk)
plot.frequency.spectrum(Zk[1:20000])
```

```{r}
dynspec(z, wl=1024, osc=T)  # function seewave
```

```{r spectrogram}
spectro(z)  # give spectrogram
```

```{r}
meanspec(z)  # obtain average for the whole time length
```

```{r}

```
